Expos\'{e} Master Thesis\\ Checking Assertions for Smart Contracts

### 2020-10-04 meeting-notes/20200928-expose.tex

\begin{itemize}
\item Formally define a set of logical formulas that are amenable to the kind of checking
  described above. In a first approximation, the interesting assertion formulas seem to
  be those with prenex universal quantification. 
\item Collect examples for interesting formulas.
\item Formalize the cost incurred by checking a formula.
\item Extend the contract language with assertions stating these logical formulas.
\item Implement the transformation: negation of the formula followed by transformation
  of existentials to picking a random value. For initial experiments, the target of the
  translation can be any language; later on, the target has to be a language hosted on
  the blockchain.

  Negation turns prenex universial quantification into prenex existential quantification. But it
  would be ineffective to generate a random integer and then check that it falls in the
  interesting range (e.g., $2\le n \le \sqrt p$ and $0\le k< n-1$). Rather, the formula
  needs to be analzed and the random generator should be restricted to only generate
  interesting values.

@@

\begin{itemize}
\item   Once a counterexample has been found,
  this counterexample is broadcast to the network where it can be verified by further
  nodes.  I think this broadcast and the reaction to this broadcast requires an update of
  the protocol: all receivers of the broadcast will check the counterexample and revert
  the transaction. 
\item A transaction using counterexamples has to
  wait for a couple of cycles in the transaction pool until the node feels safe to
  include the transaction in a block.
  The number of cycles to wait depends on the contract and on the size of the input.
  So that number should be variable.
\item (low priority) This approach poses the danger of a DOS attack. Someone could post a false
  counterexample, which does not really lead to a failure. If everyone switches to
  check this proposed counterexample instead of generating a random number, then the
  real problem may remain hidden. So, there has to be a component in the protocol that
  penalizes someone who broadcasts a false counterexample. Moreover, not all validators
  should switch to recheck a proposed counterexample; or they could check it in
  addition to their own random check.

>>>>

\begin{itemize}
\item Formally define a set of logical formulas that are amenable to the kind of checking
  described above. In a first approximation, the interesting assertion formulas seem to
  be those with prenex universal quantification. 
\item Collect examples for interesting formulas.
\item Formalize the cost incurred by checking a formula.
\item Extend the contract language with assertions stating these logical
  formulas. There are several approaches to perform this extension.
  \begin{enumerate}
  \item (Solidity) Extend the syntax of the assert statement. Requires an extension of
    the compiler.
  \item (Tezos/Solidity) The assertion is stated in its own syntax in a file separate from the
    contract implementation. Requires a dedicated compiler of assertions to
    Michelson. The generated code is either combined with the contract implementation
    or stored separately on the blockchain. Several entry points may lead to difficulties.
  \end{enumerate}
\item Implement the transformation in two steps:
  \begin{enumerate}
  \item negation of the formula
  \item followed by the translation of an existentially quantified formula to code that picks a
    random value subject to predicate. For initial experiments, the target of the
    translation can be any programming 
    language; later on, the target has to be a language hosted on the blockchain
    (e.g. Michelson or Solidity). These languages have to be extended with a random
    generator. There has to be a special execution mode for ``challenge contracts''
    that search for a counterexample of an assertion.
  \end{enumerate}
  Negation turns prenex universial quantification into prenex existential quantification. But it
  would be ineffective to generate a random integer and then check that it falls in the
  interesting range (e.g., $2\le n \le \sqrt p$ and $0\le k< n-1$). Rather, the formula
  needs to be analzed and the random generator should be restricted to only generate
  interesting values.

@@

\begin{itemize}
\item (Tezos) Proposal for an architecture. The assertion for contract C gets compiled
  to a separate contract A(C). A(C) takes C's argument x and attempts to challenge 
  the assertion by finding a counterexample. Given x, we determine a number n of
  ``probation cycles'' during which the net attempts to find a counterexample. The
  injecting node broadcasts A(C)(x) for n cycles. If A(C)(x) finishes successfully,
  this execution vanishes (as no counterexample has been found). If A(C)(x) finds a
  counterexample, it records the counterexample along with the data to check it on the
  blockchain as in A(C,x,y) where it can be validated by the net as usual.
  After n cycles the injecting node broadcasts C(x), which is processed as
  usual. However, a valid blockchain either contains C(x) or A(C,x,y), but not both. It
  is an error to attempt to enter C(x) after A(C,x,y) is on the chain, and vice versa.

  It needs to be investigated which aspects of the protocol need to be
  adapted. Moreover, A(C,x) must execute in a different mode that enables the random
  instruction. This instruction is not available in the default execution mode.
\item   Once a counterexample has been found,
  this counterexample is broadcast to the network where it can be verified by further
  nodes.  I think this broadcast and the reaction to this broadcast requires an update of
  the protocol: all receivers of the broadcast will check the counterexample and revert
  the transaction. 
\item (see above for a revised approach) A transaction using counterexamples has to
  wait for a couple of cycles in the transaction pool until the node feels safe to
  include the transaction in a block.
  The number of cycles to wait depends on the contract and on the size of the input.
  So that number should be variable.
\item (low priority) This approach poses the danger of a DOS attack. Someone could post a false
  counterexample, which does not really lead to a failure. If everyone switches to
  check this proposed counterexample instead of generating a random number, then the
  real problem may remain hidden. So, there has to be a component in the protocol that
  penalizes someone who broadcasts a false counterexample. Moreover, not all validators
  should switch to recheck a proposed counterexample; or they could check it in
  addition to their own random check.
